Big Ideas in Sports Analytics and Statistical Tools for their Investigation

### 2022-08-01 sports_analytics.Rmd
editing

These are just the most obvious examples of strategic changes that are fueled by insights extracted from data by practitioners of sports analytics.
Similar insights are now being made in less obvious settings, including esports [@clark2020bayesian; @maymin2021smart]. 
These insights come both from academia, where researchers typically use public data to produce high-caliber, peer-reviewed scientific work, as well as from industry, where highly-trained analysts work with with players, coaches, and team officials to gain immediate effect thanks to high-resolution, often proprietary data.
A growing pool of people move seamlessly between these two worlds, leading to the formation of partnerships and the cross-pollination of ideas.

@@

In many sports, the first step towards an analytical understanding of the game is the estimation of the expected value of a game state at any given point in a sporting contest. 
Mathematically, we define $X$ to be a random variable indicating the number of points (or runs) that a team will score over some determined amount of time (e.g. remainder of game, quarter, period, or inning). 
Let $s \in S$ be a tuple that encodes the *state* of a game. 
Then our task is to estimate:
$$
  \mathbb{E}[X | s] = \sum_{X \geq 0} \Pr[X | s] \cdot X\,,
$$
for any state $s \in S$, where $\Pr[X | s]$ is the probability of scoring $X$ points given that the game is in state $s$ and $S$ is the set of all possible states.

@@

erm |>
  pivot_wider(names_from = Out, values_from = Runs) |>
  mutate(Base = int2bin(Base)) |>
  kable(booktabs = TRUE, 
        linesep = "",
        caption = "George Lindsey's expected run matrix. Note how (when reading across the rows) the expected runs decrease as outs increase for the same configuration of baserunners, while (when reading down the columns) expected runs generally increase as baserunners advance. \\label{tab:erm}") |> 
  add_header_above(c("", "Out" = 3))
```

@@

First, the values within the expected run matrix change over time. 
For example, any estimation of the values in the expected run matrix based on data from a high-scoring era (e.g., the early 2000s) will yield different values than equivalent analysis in a low-scoring era. 
In a high run-scoring environment, where there are many home runs, the value of a walk may be higher, since a player who walks is more likely to score on a subsequent home run. 
Conversely, in a low run-scoring environment, stolen bases and sacrifice bunts may be comparatively more valuable. 
Thus, a careful estimate of $\mathbb{E}[X | s]$ would include a time parameter $t$, indicating when the estimate is appropriate.

@@

The first effort of estimating expected points of possession in football goes back to @Carter1971operations. 
In this technical note, the expected points for 1st and 10 plays in the NFL are estimated, given any yard line on the football field. 
Due to limitations regarding the amount of data collected, the authors divide football field into 10-yard buckets, centered at their midpoints (e.g. 5, 15, 25, 35, etc.), before averaging the value of the next scoring instance across the field to obtain the expected points.

Early work on expected point values in American football can also be found in @Carroll1988hidden.
In particular, the authors extend @Carter1971operations's work and propose a linear model for expected points in the NFL.
It is determined that every extra 25 yards is associated with 2 more points scored on average for a football team.

Another attempt at modeling expected points in football is done by @goldner2012markov and @goldner2017situational, which proposes a Markov framework.
In particular, the author considers a football drive as an *absorbing Markov chain*, consisting of distinct *absorbing states* touchdowns, field goals, and other play outcomes. 
For any given play, the expected points are calculated using the absorption probabilities for different scoring events.

A more in-depth overview of the history of expected points in sports is provided in @Yurko2019nflwar (Section 1.1). 
Most importantly, this paper uses publicly available data provided by the **nflscrapR** package [@R-nflscrapR] to model the expected points on a play-by-play level in football.
The authors introduce a multinomial logistic regression approach, which takes into account the current down, time remaining, yards from endzone, yards to go, and the indicators for goal down situation and whether there are less than two minutes remaining in the half. 
This model estimates the probabilities of the following possible scoring outcomes after each play: no score, safety, field goal, and touchdown for both the offensive and defensive teams, all of which have a point value.
The expected points for a play can then be calculated accordingly, by summing up the products of the scoring event point values and their associated probabilities.

In addition, @pelechrinis2019 develops an expected points framework in the same spirit as previous work, but accounts for the strength of the opponents in their method.
They state that by failing to account for opponent strength appropriately, about 124.8 points per team each season (or about 3.8 wins per season) are not credited correctly.

@@

@Romer2006 examines 4th down decisions in the NFL using expected points by focusing only on examples from the first quarter of a game (to avoid issues with end-of-half and end-of-game decision making). 
The author concludes that teams don't go for it enough if they are trying to maximize their probability of winning the game.

@@

Recently, @lopez2020bigger presented an introduction to NFL tracking data, and as an example, they examine 4th down behavior.
Rather than looking at the traditional distance to go, which is rounded, they have the exact position of the ball leading to an exact distance to go.
They demonstrate that teams are still conservative when choosing to go for it or punt on 4th down, but the effect may be overstated.

@@

As for quarterback evaluation, @White2002 present a tiered logistic regression method that can be, in general, applied to any regression setting with a polychotomous response.
Using this technique, they estimate the value of NFL plays using a simple expected points model with down, yards to go, and yards to goal as predictors.
Accordingly, the model results are utilized to obtain ratings and rankings for NFL passers.

In addition, @Alamar2010 implements an expected points framework to examine play calling in the NFL. 
However, rather than assessing each play on its own, the author evaluates the play in the context of the drive. 
Based on play-by-play data from 2005 through 2008, it is determined that teams are under-utilizing passing plays in some situations.

@@

See also @bornn2017studying for more information on how tracking data have enabled advanced statistical analyses of basketball in recent years.

In American football, @Yurko2020going use tracking data provided by the [2019 NFL Big Data Bowl](https://operations.nfl.com/gameday/analytics/big-data-bowl/past-big-data-bowl-recaps/) to model the expected yards gained for a ball-carrier during the course of a play. 
As an extension to pre-existing approaches, the authors use conditional density estimation to obtain a probability distribution for the number of yards gained during the play, rather than only producing a single estimate for the expected yards gained. 
Accordingly, the probability of various types of outcomes at the end of a play such as a touchdown or a first-down gain can be computed from the distribution of the end-of-play yard line.

@@

For example, in the bottom of the ninth inning of a tied baseball game, the optimal strategy for winning the game is maximizing the probability of scoring *at least one run*, which may differ from the strategy of maximizing expected runs. 
If we let $U$ be the set of all strategies, then we assert that it is not always the case that the strategy $u$ that maximizes the expected number of points will maximize the probability of winning:
$$
  \underset{u \in U}{\arg \max \,}{\Pr[X > 0 | s, u]} \neq \underset{u \in U}{\arg \max \,}{\mathbb{E}[X | s, u]} \,.
$$
Consider the situation where runners are on first and third base, and the score is tied in the bottom of the ninth inning with no one out. 
Table \ref{tab:tied} reveals that the expected number of runs scored in the remainder of the inning is 1.94 runs, while the probability of scoring zero runs is 0.13. 
The defense is in a tight spot, facing an 87% probability of losing the game. 
However, by walking the hitter to load the bases, they create the opportunity to force the lead runner at home and thus reduce the chances of scoring to 82%, even though they raise the expected number of runs scored to 2.22. 
In this case, the defensive team is wise to pursue the strategy of maximizing the expected number of runs scored, because it *minimizes* the probability of scoring at least one run.

@@

erm |>
  filter(Out == 0, Base %in% c(5, 7)) |>
  mutate(Base = int2bin(Base)) |>
  kable(booktabs = TRUE, 
        caption = "Loading the bases can be a rational strategy. In this case, loading the bases maximizes both the expected number of runs scored *and* the probability of not scoring at all. \\label{tab:tied}")
```

@@

The notion of win probability in baseball goes back to at least as early as @Lindsey1961.
In this paper, the author calculates the expected win probability after each inning based on the distribution of runs scored in each inning. 
Inspired by @Lindsey1961's work, @Mills1970 utilize win probability to introduce Player Win Average (PWA), a measure of a player's contribution to the game outcome. 
In particular, PWA is computed as 
$$
PWA = \frac{Win \ Points}{Win \ Points + Loss \ Points},
$$
where the win and loss points represent how much the player positively and negatively impacts their team's probability of winning after each play.
In effect, the win points are the sum of the changes in $\Pr[W_i | s]$ from one state to the next.

@@

A third crucial idea in sports analytics is the estimation of team strength. 
A popular method for estimating team strength in sports is through pairwise evaluations. 
Perhaps the most famous probability model for predicting the outcome of a paired comparison is the Bradley-Terry model (BTM) [@Bradley1952]. 
For a pair of players (or teams) $i$ and $j$, let $\Pi_{ij}$ denote the probability that $i$ is preferred to $j$. 
Then the BTM is a logistic regression model with parameters $\beta_i, \beta_j$ such that
$$
\log\left(\frac{\Pi_{ij}}{\Pi_{ji}}\right) = \beta_i - \beta_j \,.
$$
Here, $\exp{\beta_i}$ is often viewed as a representation of player $i$'s ability.

@@

Table \ref{tab:tennisbtm} shows the estimated coefficients of the fitted BTM. 
According to the abilities, between 2014 and 2018 the players are ranked as follows: 1) Djokovic, 2) Federer, 3) Wawrinka, 4) Nadal, 5) Murray.
To obtain win probabilities, as an illustration, for the Federer-Nadal matchup, an estimate for the probability of a Federer victory is
$$
\hat\Pi_{24} = \frac{exp(\hat\beta_2 - \hat\beta_4)}{1 + exp(\hat\beta_2 - \hat\beta_4)} = \frac{exp(1.136 + 0.062)}{1 + exp(1.136 + 0.062)} = 0.768.
$$

@@

Another widely known tool for measuring team strength is the Elo rating system [@Elo1978]. The Elo rating was originally developed for chess, but also has been used to estimate team strength in other sports. 
See @koning2017rating for more information on applications of the Elo rating in soccer.

@Glickman1998 propose a Bayesian state-space model for paired comparisons for predicting NFL games, allowing team strength parameters to vary over time. 
In particular, they model point differential in the NFL by introducing week-to-week and season-to-season as the two primary sources of variation in team strengths. 
See also @glickman2017estimating for more discussion on estimating team strengths in American football. 
More recently, @Lopez2018 extend @Glickman1998's state-space model to understand randomness in the four major American sports leagues. 
Betting moneylines are used in place of point differentials in order to estimate team strengths, and this framework also accounts for home advantage.

@@

```{r chess, eval=whoami::username() == 'bbaumer', fig.show="hold", out.width="49%", fig.cap="At left, the starting chess board printed via the \\pkg{chess} package. At right, the final position for one of the authors' recent wins (a checkmate playing Black)."}
library(chess)
# Sys.setenv(RETICULATE_PYTHON = "my_env/bin/python3")
game() |> plot()

@@

## Colors

Each professional sports team has its own brand, most obviously identified by a team logo and set of colors. 
The \pkg{teamcolors} R package [@R-teamcolors] provides color palettes and logos for men's and women's professional and collegiate sports teams, as well as color and fill scale functions compatible with \pkg{ggplot2} [@R-ggplot2].
For instance, the NFL teams' colors and logos shown in Figure \ref{fig:nflfastr-wp} were provided by the \pkg{teamcolors} package. 
Figure \ref{fig:teamcolors} illustrates the full array of team color palettes for professional sports leagues.

@@

# Opportunities

@@

Perhaps motivated by his success in the Kaggle March Madness competition, Michael Lopez joined the National Football League and launched the [Big Data Bowl](https://operations.nfl.com/gameday/analytics/big-data-bowl/) in 2019.
This annual competition has similarly fueled new research directions in American football and a [*JQAS* special issue on player tracking data in the National Football League (NFL)](https://www.degruyter.com/journal/key/jqas/16/2/html) of [@lopez2020bigger].
Successful entries and their corresponding publications [@DeshpandeEvans2020; @Yurko2020going; @ChuReyersThomsonWu2020] have launched the careers of several of the most prominent early-career researchers in sports analytics.

@@

As an applied science, sports analytics may lack a [grand unified theory](https://en.wikipedia.org/wiki/Grand_Unified_Theory) that succinctly characterizes game play across all sports. 
However, as a maturing discipline, sports analytics has been able to address fundamental questions common to many sports. 
In this paper, we explicate three of those big questions:

@@

# Acknowledgments

>>>>

These are just the most obvious examples of strategic changes that are fueled by insights extracted from data by practitioners of sports analytics.
Similar insights are now being made in less obvious settings, including esports [@clark2020bayesian; @maymin2021smart]. 
These insights come both from academia, where researchers typically use public data to produce high-caliber, peer-reviewed scientific work, as well as from industry, where highly-trained analysts work with with players, coaches, and team officials to put new ideas into immediate effect thanks to high-resolution, often proprietary data.
A growing pool of people move seamlessly between these two worlds, leading to the formation of partnerships and the cross-pollination of ideas.

@@

In many sports, the first step towards an analytical understanding is the estimation of the expected value of a game state at any given point in it. 
Mathematically, we define $X$ to be a random variable indicating the number of points (or runs) that a team will score over some determined amount of time (e.g., remainder of game, quarter, period, or inning). 
Let $s \in S$ be a tuple that encodes the *state* of a game. 
Then our task is to estimate:

\begin{equation}
  \mathbb{E}[X | s] = \sum_{X \geq 0} \Pr[X | s] \cdot X\,,
\end{equation}

for any state $s \in S$, where $\Pr[X | s]$ is the probability of scoring $X$ points given that the game is in state $s$ and $S$ is the set of all possible states.

@@

erm |>
  pivot_wider(names_from = Out, values_from = Runs) |>
  mutate(Base = int2bin(Base)) |>
  kable(booktabs = TRUE, 
        linesep = "",
        caption = "George Lindsey's expected run matrix. Note how (when reading across the rows) the expected runs decrease as outs increase for the same configuration of baserunners, while (when reading down the columns) expected runs generally increase as baserunners advance. 000 means no runners on base, and 110 means runners on second and third bases. \\label{tab:erm}") |> 
  add_header_above(c("", "Out" = 3))
```

@@

First, the values within the expected run matrix change over time. 
For example, any estimation of the values in the expected run matrix based on data from a high-scoring era (e.g., the early 2000s) will yield different values than equivalent analysis in a low-scoring era. 
In a high run-scoring environment, where there are many home runs, the value of a walk may be higher, since a player who walks is more likely to score on a subsequent home run. 
Conversely, in a low run-scoring environment where hits are hard to come by, stolen bases and sacrifice bunts may be comparatively more valuable. 
Thus, a careful estimate of $\mathbb{E}[X | s]$ would include a time parameter $t$, indicating when the estimate is appropriate.

@@

The task of estimating expected points of possession in football goes back to @Carter1971operations, who estimate the expected points for 1st and 10 plays in the NFL, given any yard line on the football field. 
Due to limitations regarding the amount of data collected, the authors divide football field into 10-yard buckets, centered at their midpoints (e.g. 5, 15, 25, 35, etc.), before averaging the value of the next scoring instance across the field to obtain the expected points.

XX need example.

Early work on expected point values in American football can also be found in @Carroll1988hidden.
In particular, the authors extend @Carter1971operations's work and propose a linear model for expected points in the NFL.
They determine that every extra 25 yards is associated with 2 more points scored on average for a football team.

Other attempts at modeling expected points in football are @goldner2012markov and @goldner2017situational, who propose a Markov framework.
In particular, the author considers a football drive as an *absorbing Markov chain*, consisting of distinct *absorbing states* touchdowns, field goals, and other play outcomes. 
For any given play, the expected points are calculated using the absorption probabilities for different scoring events.

A more in-depth overview of the history of expected points in sports is provided in @Yurko2019nflwar (Section 1.1). 
Most importantly, @Yurko2019nflwar use publicly available data provided by the **nflscrapR** package [@R-nflscrapR] to model the expected points on a play-by-play level in football.
The authors introduce a multinomial logistic regression approach, which takes into account the current down, time remaining, yards from endzone, yards to go, and indicators for goal down situation and whether there are less than two minutes remaining in the half. 
Their model estimates the probabilities of the following possible scoring outcomes after each play: no score, safety, field goal, and touchdown for both the offensive and defensive teams, all of which have a point value.
The expected points for a play can then be calculated accordingly, by summing up the products of the scoring event point values and their associated probabilities (see Equation 1).

In addition, @pelechrinis2019 develop an expected points framework in the same spirit as the previous work, but account for the strength of the opponents in their method.
They state that by failing to account for opponent strength appropriately, about 124.8 points per team each season (or about 3.8 wins per season) are not credited correctly.

XX Seems massive

@@

@Romer2006 examines 4th down decisions in the NFL using expected points by focusing only on examples from the first quarter of a game (to avoid issues with end-of-half and end-of-game decision making). 
He concludes that teams don't go for it enough if they are trying to maximize their probability of winning the game.

@@

Recently, @lopez2020bigger presents an introduction to NFL tracking data, and as an example, examines 4th down behavior.
Rather than looking at the traditional distance to go, which is rounded, they have the exact position of the ball leading to an exact distance to go.
They demonstrate that teams are still conservative when choosing to go for it or punting on 4th down, but the effect may be overstated.

@@

For quarterback evaluation, @White2002 present a tiered logistic regression method that can be, in general, applied to any regression setting with a polychotomous response.
Using this technique, they estimate the value of NFL plays using a simple expected points model with down, yards to go, and yards to goal as predictors.
Accordingly, the model results are utilized to obtain ratings and rankings for NFL passers.

In addition, @Alamar2010 implements an expected points framework to examine play calling in the NFL. 
However, rather than assessing each play on its own, the author evaluates the play in the context of the drive. 
Based on play-by-play data from 2005 through 2008, he determines that teams are under-utilizing passing plays in some situations.

@@

See also @bornn2017studying for more information on how tracking data have enabled advanced statistical analyses of basketball in recent years.
The strategy of maximizing expected points in basketball has lead directly to the proliferation of three-point shooting in the NBA.

In American football, @Yurko2020going use tracking data provided by the [2019 NFL Big Data Bowl](https://operations.nfl.com/gameday/analytics/big-data-bowl/past-big-data-bowl-recaps/) (see Section \ref{sec:opportunities}) to model the expected yards gained for a ball-carrier during the course of a play. 
As an extension to pre-existing approaches, the authors use conditional density estimation to obtain a probability distribution for the number of yards gained during the play, rather than only producing a single estimate for the expected yards gained. 
Accordingly, the probability of various types of outcomes at the end of a play such as a touchdown or a first-down gain can be computed from the distribution of the end-of-play yard line.

@@

For example, in the bottom of the ninth inning of a tied baseball game, the optimal strategy for winning the game is maximizing the probability of scoring *at least one run*, which may differ from the strategy of maximizing expected runs. 
If we let $U$ be the set of all strategies, then we assert that it is not always the case that the strategy $u \in U$ that maximizes the expected number of points will maximize the probability of winning:
$$
  \underset{u \in U}{\arg \max \,}{\Pr[X > 0 | s, u]} \neq \underset{u \in U}{\arg \max \,}{\mathbb{E}[X | s, u]} \,.
$$
Consider the situation where runners are on first and third base, and the score is tied in the bottom of the ninth inning with no one out. 
Table \ref{tab:tied} (derived from Table \ref{tab:erm}) reveals that the expected number of runs scored in the remainder of the inning is 1.94 runs, while the probability of scoring zero runs is 0.13. 
The defense is in a tight spot, facing an 87% probability of losing the game. 
However, by walking the hitter to load the bases, they create the opportunity to force the lead runner at home and thus reduce the chances of scoring to 82%, even though they raise the expected number of runs scored to 2.22. 
In this case, the *defensive* team is wise to pursue the strategy of maximizing the expected number of runs scored, because it *minimizes* the probability of scoring at least one run.

@@

erm |>
  filter(Out == 0, Base %in% c(5, 7)) |>
  mutate(Base = int2bin(Base)) |>
  rename(`Pr(Scoring)` = Pr_0) |>
  kable(
    booktabs = TRUE, 
    caption = "Loading the bases can be a rational strategy. In this case, loading the bases maximizes both the expected number of runs scored and the probability of not scoring at all. \\label{tab:tied}"
  )
```

@@

The notion of win probability in baseball goes back to at least as early as @Lindsey1961, who calculates the expected win probability after each inning based on the distribution of runs scored in each inning. 
Inspired by @Lindsey1961's work, @Mills1970 utilize win probability to introduce Player Win Average (PWA), a measure of a player's contribution to the game outcome. 
In particular, PWA is computed as 
$$
PWA = \frac{Win \ Points}{Win \ Points + Loss \ Points},
$$
where the win and loss points represent how much the player positively and negatively impacts their team's probability of winning after each play.
In effect, the win points are the sum of the changes in $\Pr[W_i | s]$ from one state to the next.

@@

A third crucial idea in sports analytics is the estimation of team strength. 
A popular method for estimating team strength in sports is through pairwise evaluations. 
Perhaps the most famous probability model for predicting the outcome of a paired comparison is the Bradley-Terry model (BTM) [@Bradley1952]. 
For a pair of players (or teams) $i$ and $j$, let $\Pi_{ij}$ denote the probability that $i$ is preferred to $j$. 
Then the BTM is a logistic regression model with parameters $\beta_i, \beta_j$ such that
$$
\log\left(\frac{\Pi_{ij}}{\Pi_{ji}}\right) = \beta_i - \beta_j \,.
$$
Here, $\exp{\beta_i}$ is often viewed as a representation of team $i$'s ability.

@@

Table \ref{tab:tennisbtm} shows the estimated coefficients of the fitted BTM. 
According to the abilities, between 2014 and 2018 the players are ranked as follows: 1) Djokovic, 2) Federer, 3) Wawrinka, 4) Nadal, 5) Murray.
In addition to an ordering, the magnitude of the coefficients in Table \ref{tab:tennisbtm} provide a measure of relative strength.

To obtain win probabilities, as an illustration, for the Federer-Nadal matchup, an estimate for the probability of a Federer victory is:
$$
\hat\Pi_{24} = \frac{exp(\hat\beta_2 - \hat\beta_4)}{1 + exp(\hat\beta_2 - \hat\beta_4)} = \frac{exp(1.136 + 0.062)}{1 + exp(1.136 + 0.062)} = 0.768 \,.
$$

@@

Another widely known tool for measuring team strength is the Elo rating system [@Elo1978], which was originally developed for chess, but also has been used to estimate team strength in other sports. 
See @koning2017rating for more information on applications of the Elo rating in soccer.

@Glickman1998 propose a Bayesian state-space model for paired comparisons for predicting NFL games, allowing team strength parameters to vary over time. 
In particular, they model point differential in the NFL by introducing week-to-week and season-to-season as the two primary sources of variation in team strengths. 
See also @glickman2017estimating for more discussion on estimating team strengths in American football. 
More recently, @Lopez2018 extend @Glickman1998's state-space model to understand randomness in the four major American sports leagues. 
Betting moneylines are used in place of point differentials in order to estimate team strengths, and this framework also accounts for home advantage.
Both papers motivate the usefulness of model-based measures of team strength by demonstrating their superiority to low-resolution win-loss records. 
Apart from sports gambling, having an accurate estimate of team strength is useful to team officials, who are constantly monitoring and forecasting their team's ability.

@@

```{r chess, eval=whoami::username() == 'bbsdaumer', fig.show="hold", out.width="49%", fig.cap="At left, the starting chess board printed via the \\pkg{chess} package. At right, the final position for one of the authors' recent wins (a checkmate playing Black)."}
library(chess)
# Sys.setenv(RETICULATE_PYTHON = "my_env/bin/python3")
game() |> plot()

@@

## Theming

Each professional sports team has its own brand, most obviously identified by a team logo and set of colors. 
The \pkg{teamcolors} R package [@R-teamcolors] provides color palettes and logos for men's and women's professional and collegiate sports teams, as well as color and fill scale functions compatible with \pkg{ggplot2} [@R-ggplot2].
For example, the NFL teams' colors and logos shown in Figure \ref{fig:nflfastr-wp} were provided by the \pkg{teamcolors} package. 
Figure \ref{fig:teamcolors} illustrates the full array of team color palettes for professional sports leagues.

@@

# Opportunities {#sec:opportunities}

@@

Perhaps motivated by his success in the Kaggle March Madness competition, Michael Lopez joined the National Football League and launched the [Big Data Bowl](https://operations.nfl.com/gameday/analytics/big-data-bowl/) in 2019.
This annual competition has similarly fueled new research directions in American football and a [*JQAS* special issue on player tracking data in the National Football League](https://www.degruyter.com/journal/key/jqas/16/2/html) [@lopez2020bigger].
Successful entries and their corresponding publications [@DeshpandeEvans2020; @Yurko2020going; @ChuReyersThomsonWu2020] have launched the careers of several of the most prominent early-career researchers in sports analytics.

@@

As an applied science, sports analytics may lack a [grand unified theory](https://en.wikipedia.org/wiki/Grand_Unified_Theory) that succinctly characterizes game play across all sports. 
However, as a maturing discipline, sports analytics has been able to address fundamental questions common to many sports. 
In this paper, we explore three of those big questions:

@@

# Acknowledgments {.unnumbered}
